Optimal (Multiway) Spatial Joins

### 2024-05-24 msj.tex
yufei

\begin{theorem} \label{thm:main-recur}
    For any $n \ge 1$, equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in time
    \myeqn{
        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big).
        %\nn
        \label{eqn:main:reccurrence}
    }
%     time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
\end{theorem}

@@

Our objective is to prove H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ. For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$. Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$. We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.
We will prove:

\begin{lemma} \label{lmm:hv}
    The H-V $k$-SJ problem can be solved in time
    \myeqn{
        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big). \nn 
    }
\end{lemma}

@@

We will find result tuples of each type separately. Before proceeding, however, let us observe a sense of symmetry for Type 2. As mentioned, for each result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both. As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc. It is possible for the result tuple to belong to multiple sub-types simultaneously.

Our subsequent discussion will proceed as follows.  Section~\ref{sec:hv:type1} will explain how to produce all the result tuples of Type 1. Section~\ref{sec:hv:type2} will explain how to produce the result tuples $(r_1, ..., r_{k-2}, h, v)$ of a particular sub-type of Type 2: those tuples where $r_1$ covers the left endpoint of $h$. Other sub-types can be computed using the same algorithm, by symmetry.

A remark is in order about duplicate removal. A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types. Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm. However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times. Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to. To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting. One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.

\subsection{Finding Result Tuples of Type 1} \label{sec:hv:type1}

@@

\subsection{Finding Result Tuples of Type 2} \label{sec:hv:type2}

>>>>

\begin{theorem} \label{thm:main-recur}
    Equipped with the algorithm $\A$ as described above, the $k$-SJ problem can be solved in time
    \myeqn{
        O(k^3) \cdot \big( F_{k-1}(n, \out) + n \log n + k \cdot \out \big)
        %\nn
        \label{eqn:main:reccurrence}
    }
    where $n$ (resp., $\out$) is the input (resp., output) size of the problem.
%     time using $O(n + \out)$ space, plus the worst-case space of $\A$ on any $(k-1)$-SJ instance with input size $n$ and output size $\out$.
\end{theorem}

@@

Our objective is to prove H-V $k$-SJ can be efficiently reduced to $(k-1)$-SJ --- note: it is $(k-1)$-SJ here, rather than H-V $(k-1)$-SJ. For this purpose, we assume the existence of an algorithm $\A$ that can solve $\kappa$-SJ problems for all $\kappa \in [2, k-1]$. Denote by $F_\kappa(n, \out)$ the worst-case runtime of $\A$ on any instance of $\kappa$-SJ that has input size $n$ and output size $\out$. We consider that $F_\kappa(n, \out) \le F_{\kappa + 1}(n, \out)$, that is, its overhead on $\kappa$-SJ should not be larger than that on $(\kappa+1)$-SJ.

We will establish:

\begin{lemma} \label{lmm:hv}
    The H-V $k$-SJ problem can be solved in time
    \myeqn{
        O(k) \cdot \big( F_{k-1}(n, \out) + n\log n + k \cdot \out \big) \nn
    }
    where $n$ (resp., $\out$) is the input (resp., output) size of the problem.
\end{lemma}

@@

In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.

%We will find result tuples of each type separately. Before proceeding, however, let us observe a sense of symmetry for Type 2. As mentioned, for each result tuple $(r_1, ..., r_{k-2}, h, v)$ of this type, a rectangle $r_i$, for some $i \in [k-2]$, covers an endpoint of $h$ or $v$ or both. As (i) there are $k-2$ choices for $i$ and (ii) $h$ and $v$ together have four endpoints, we can divide Type 2 further into $4(k-2)$ ``sub-types'': e.g., in subtype 1 (resp., 2), $r_1$ covers the left (resp., right) endpoint of $h$, in subtype 3 (resp., 4), $r_1$ covers the bottom (resp., top) endpoint of $v$, in subtype 5 (resp., 6), $r_2$ covers the left (resp., right) endpoint of $h$, etc. It is possible for the result tuple to belong to multiple sub-types simultaneously.

\vgap 
%
% Our subsequent discussion will proceed as follows.  Section~\ref{sec:hv:type1} will explain how to produce all the result tuples of Type 1. Section~\ref{sec:hv:type2} will explain how to produce the result tuples $(r_1, ..., r_{k-2}, h, v)$ of a particular sub-type of Type 2: those tuples where $r_1$ covers the left endpoint of $h$. Other sub-types can be computed using the same algorithm, by symmetry.
%
% \vgap
%
% A remark is in order about duplicate removal. A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types. Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm. However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times. Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to. To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting. One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.

\section{H-V $k$-SJ: Result Tuples of Type 1} \label{sec:hv:type1}

@@

\section{Finding Result Tuples of Type 2} \label{sec:hv:type2}
